---
title: "HbA1c Case"
author: "Anders Ellegaard"
date: "2023 01 03"
output:
  ioslides_presentation:
    widescreen: yes
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE)
library(magrittr)
library(ggplot2)
colour <- list(
  orange_dark = "#fb8500",
  orange_light = "#ffb703",
  blue_dark = "#023047",
  azure = "#219ebc",
  blue_light = "#8ecae6"
)
```

## Contents

 - Code for delivery 1
 - Delivery 1: table
 - Code for delivery 2
 - Delivery 2: plot
 - Extra: Average treatment difference

```{r duplicates, include = FALSE}
# While preparing for the table, I spotted a duplicate
readxl::read_excel("data.xlsx", sheet = "ADLB") %>%
  dplyr::filter(VISITNUM == 10, PARAMCD == "C64849B") %>%
  dplyr::group_by(SUBJID) %>%
  dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
  dplyr::filter(n > 1)
```

## Data Prep for Table {.smaller}

```{r echo=TRUE}
hba1c <- readxl::read_excel("data.xlsx", sheet = "ADLB") %>%
  dplyr::filter(VISITNUM == 10, PARAMCD == "C64849B", ANL01FL == "Y") %>%
  dplyr::select(SUBJID, AVAL) %>%
  dplyr::rename(HbA1c = AVAL)
# Just filter, join, and summarise
tbl <- readxl::read_excel("data.xlsx", sheet = "ADSL") %>%
  dplyr::filter(FASFL == "Y") %>%
  dplyr::left_join(hba1c, by = "SUBJID") %>%
  dplyr::bind_rows(., dplyr::mutate(., TRT01P = "Total")) %>%
  dplyr::group_by(TRT01P) %>%
  dplyr::summarise(dplyr::across(
    where(is.double),
    list(
      N_nna = ~ sum(!is.na(.x)),
      N_na = ~ sum(is.na(.x)),
      mean = ~ mean(.x, na.rm = TRUE),
      sd = ~ sd(.x, na.rm = TRUE),
      median = ~ median(.x, na.rm = TRUE),
      min = ~ min(.x, na.rm = TRUE),
      max = ~ max(.x, na.rm = TRUE)
    ),
    .names = "{.col}-{.fn}"
  )) %>%
  tidyr::pivot_longer(-TRT01P) %>%
  tidyr::separate(name, into = c("param", "fun"), sep = "-") %>%
  tidyr::pivot_wider(
    id_cols = c(param, fun),
    names_from = TRT01P,
    values_from = value
  )
```

## Delivery 1: Table

```{r}
tbl %>%
  dplyr::rename(
    Parameter = param,
    Aggregate = fun
  ) %>%
  DT::datatable(
    options = list(
      pageLength = 7,
      dom = "tlp",
      lengthMenu = c(7, 14, 21, 28)
    ),
    rownames = FALSE
  ) %>%
  DT::formatSignif(
    columns = c("Total", "Treatment A", "Treatment B"),
    digits = 3
  )
```

## Data Prep for Plot {.smaller}

```{r echo=TRUE}
hba1c <- readxl::read_excel("data.xlsx", sheet = "ADLB") %>%
  dplyr::filter(FASFL == "Y", PARAMCD == "C64849B", ANL01FL == "Y") %>%
  dplyr::rename(HbA1c = AVAL) %>%
  dplyr::mutate(
    week = stringr::str_extract(AVISIT, "Week (\\d+)", group = 1),
    week = as.numeric(week)
  )
plt <- hba1c %>%
  dplyr::group_by(TRTP, week) %>%
  dplyr::summarise(
    median = median(HbA1c, na.rm = TRUE),
    q55 = quantile(HbA1c, 0.055, na.rm = TRUE),
    q945 = quantile(HbA1c, 0.945, na.rm = TRUE),
    .groups = "drop"
  )
```

## Delivery 2: Plot {.flexbox .vcenter .smaller}

```{r plot}
integers_between <- function(a, b) {
  l <- ceiling(min(a, b))
  u <- floor(max(a, b))
  if (l > u) return()
  return(seq.int(l, u, 1))
}
ymin <- 5
ymax <- 10
ggplot(plt, aes(x = week, y = median, colour = TRTP)) +
  geom_hline(yintercept = 6.5, alpha = 0.5) +
  geom_hline(yintercept = 7, alpha = 0.5) +
  geom_errorbar(
    aes(ymin = q55, ymax = q945),
    alpha = 0.5,
    width = 1.5,
    linetype = 2
  ) +
  geom_point() +
  labs(x = "Week", y = "HbA1c (%)", colour = "") +
  scale_x_continuous(breaks = unique(plt[["week"]])) +
  scale_y_continuous(breaks = integers_between(ymin*2, ymax*2)/2) +
  scale_colour_manual(values = c(
    "Treatment A" = colour$blue_dark,
    "Treatment B" = colour$orange_dark
  )) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(colour = "grey90"),
    panel.grid.major.x = element_blank(),
    panel.background = element_rect(fill = "#f5f5f0"),
    plot.background = element_rect(fill = "#f5f5f0"),
    legend.background = element_rect(fill = "transparent"),
    legend.key = element_rect(fill = "transparent")
  )
```

*Points are the median HbA1c (%) for each treatment. The vertical bars span the 5.5% to 94.5% quantile range.*
*The emphasised horizontal lines represent 6.5% and 7.5% HbA1c respectively.*

# Extra

## Treatment Difference Histogram {.flexbox .vcenter .smaller}

```{r}
a <- dplyr::filter(hba1c, week == 20, TRTP == "Treatment A")
b <- dplyr::filter(hba1c, week == 20, TRTP == "Treatment B")
ggplot() +
  geom_histogram(
    data = a,
    mapping = aes(x = HbA1c, fill = TRTP),
    binwidth = 0.1,
    alpha = 0.5
  ) +
  geom_histogram(
    data = b,
    mapping = aes(x = HbA1c, fill = TRTP),
    binwidth = 0.1,
    alpha = 0.5
  ) +
  labs(y = "Count", x = "HbA1c (%)", fill = "", title = "HbA1c Week 20") +
  scale_fill_manual(values = c(
    "Treatment A" = colour$blue_dark,
    "Treatment B" = colour$orange_dark
  )) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(colour = "grey90"),
    panel.grid.major.x = element_blank(),
    panel.background = element_rect(fill = "#f5f5f0"),
    plot.background = element_rect(fill = "#f5f5f0"),
    legend.background = element_rect(fill = "transparent"),
    legend.key = element_rect(fill = "transparent")
  )
```

## Average Treatment Difference Model {.smaller}

```{r model, cache=TRUE, results=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
treatments <- unique(hba1c$TRTP)
weeks <- unique(hba1c$week)
data <- list(
  N = nrow(hba1c),
  N_trt = length(treatments),
  N_week = length(weeks),
  trt_ind = match(hba1c$TRTP, treatments),
  week_ind = match(hba1c$week, weeks),
  y = hba1c$HbA1c
)
post <- rstan::stan(
  "trt_diff.stan",
  data = data,
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 4444
)
res <- rstan::extract(post) %>%
  tibble::as_tibble() %>%
  dplyr::pull(mu_diff)
samples <- purrr::imap(weeks, function(w, i) {
  tibble::tibble(week = w, mu_diff = res[ , i])
}) %>%
  dplyr::bind_rows()
```


## Average Treatment Difference Week 20 {.flexbox .vcenter .smaller}

```{r}
samples %>%
  dplyr::filter(week == 20) %>%
  ggplot(aes(x = mu_diff)) +
  geom_histogram(alpha = 0.5, fill = colour$orange_dark, binwidth = 0.01) +
  labs(y = "", x = "Difference in HbA1c (%) B - A", fill = "", title = "B - A Average Treatment Difference, Week 20") +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(colour = "grey90"),
    panel.grid.major.x = element_blank(),
    panel.background = element_rect(fill = "#f5f5f0"),
    plot.background = element_rect(fill = "#f5f5f0"),
    legend.background = element_rect(fill = "transparent"),
    legend.key = element_rect(fill = "transparent")
  )
```


## Average Treatment Difference Summary {.flexbox .vcenter .smaller}

```{r}
samples_summary <- samples %>%
  dplyr::group_by(week) %>%
  dplyr::summarise(
    median = median(mu_diff, na.rm = TRUE),
    q55 = quantile(mu_diff, 0.055, na.rm = TRUE),
    q945 = quantile(mu_diff, 0.945, na.rm = TRUE),
    .groups = "drop"
  )

ymin <- -1
ymax <- 1
ggplot(samples_summary, aes(x = week, y = median)) +
  geom_errorbar(
    aes(ymin = q55, ymax = q945),
    alpha = 0.5,
    width = 1.5,
    linetype = 2,
    colour = colour$orange_dark
  ) +
  geom_point(colour = colour$orange_dark) +
  labs(x = "Week", y = "Difference HbA1c (%)") +
  scale_x_continuous(breaks = unique(samples_summary[["week"]])) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(colour = "grey90"),
    panel.grid.major.x = element_blank(),
    panel.background = element_rect(fill = "#f5f5f0"),
    plot.background = element_rect(fill = "#f5f5f0"),
    legend.background = element_rect(fill = "transparent"),
    legend.key = element_rect(fill = "transparent")
  )
```

*Points are the median treatment difference (B- A) in HbA1c (%). The vertical bars span the 5.5% to 94.5% quantile range.*
